Approximation as Differentiation: Notes

### 2025-05-25 main/category-of-families.tex
spivak65.
Less noisy notation for constant family.

Derivatives respect composition. Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$. For any $x \in \RR^m$
we have:

@@

We can verify that $\tangents$ is indeed a functor, regardless of whether the linear maps $\RR^m \linearto
\RR^n$ happen to be derivatives. But if $f_2(x)$ is the derivative of $f_1$ at $x$ for any $x \in \RR^m$ and
similarly for $g$ then the derivatives compose according to the forward chain rule and $(g \comp f)_2(x)$ is
the derivative of $(g \comp f)_1$ at $x$. We can see these maps as living in $\Fam(\FinVect)$ if we regard
every $\RR^n$ as paired with a constant family of tangent spaces also of the form $\RR^n$ and each map $f:
\RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ to pairs $(f_1, f_2)$ as the pair of a differentiable function
$f_1: \RR^m \to \RR^n$ and a family of linear maps $f_2: \RR^m \to \RR^m \linearto \RR^n$.

Then for any $(f, \partial f): (\RR^m, \const(\RR^m)) \to (\RR^n, \const(\RR^n))$ and $(g,
\partial g): (\RR^n, \const(\RR^n)) \to (\RR^k, \const(\RR^k))$, we have that for any $x \in \RR^m$:
\begin{itemize}
\item $\partial f_x: \RR^m \to \RR^n$;
\item $\reindex{\partial g}{f}_x = \partial g_{f(x)}: \RR^n \to \RR^k$.
\end{itemize}

\noindent The composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
\partial f): (\RR^m, \const(\RR^m)) \to (\RR^k, \const(\RR^k))$ corresponds exactly to the forward chain rule,
since $(\reindex{\partial g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.

>>>>

Derivatives respect composition~\cite{spivak65}. Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$. For
any $x \in \RR^m$ we have:

@@

We can verify that $\tangents$ is indeed a functor, regardless of whether the linear maps $\RR^m \linearto
\RR^n$ happen to be derivatives. But if $f_2(x)$ is the derivative of $f_1$ at $x$ for any $x \in \RR^m$ and
similarly for $g$ then the derivatives compose according to the forward chain rule and $(g \comp f)_2(x)$ is
the derivative of $(g \comp f)_1$ at $x$. We can see these maps as living in $\Fam(\FinVect)$ if we regard
every finite vector space $\RR^n$ as paired with the constant family of tangent spaces which is always $\RR^n$
(also written $\RR^n$, for convenience), and each map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ to
pairs $(f_1, f_2)$ as the pair of a differentiable function $f_1: \RR^m \to \RR^n$ and a family of linear maps
$f_2: \RR^m \to \RR^m \linearto \RR^n$.

Then for any $(f, \partial f): (\RR^m, \RR^m) \to (\RR^n, \RR^n)$ and $(g, \partial g): (\RR^n, \RR^n) \to
(\RR^k, \RR^k)$, we have that for any $x \in \RR^m$:
\begin{itemize}
\item $\partial f_x: \RR^m \to \RR^n$;
\item $\reindex{\partial g}{f}_x = \partial g_{f(x)}: \RR^n \to \RR^k$.
\end{itemize}

\noindent The composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
\partial f): (\RR^m, \RR^m) \to (\RR^k, \RR^k)$ corresponds exactly to the forward chain rule,
since $(\reindex{\partial g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.

### 2025-05-25 main/higher-order.tex
Less noisy notation for constant family.

\todo{
\begin{enumerate}
\item if $\cat{C}$ has biproducts and all small products then $\Fam(\cat{C})$ is Cartesian closed
\item $\PSh(\cat{C})$ has all small products (inherited from $\Set$)
\item $\FinVect$ and $\LatGal$ have biproducts but not all small products
\item but actually we need $\PSh_{\CMon}(\cat{C})$ rather than $\PSh(\cat{C})$ because $\CMon$-enriched Yoneda embedding preserves biproducts
\end{enumerate}
}

>>>>

\todo{
\begin{enumerate}
  \item if $\cat{C}$ has biproducts and all small products then $\Fam(\cat{C})$ is Cartesian closed
  \item $\FinVect$ and $\LatGal$ both have biproducts but neither has all small products (\todo{restrict $\LatGal$ to finite lattices?})
  \item $\PSh(\cat{C})$ has all small products (inherited from $\Set$), but doesn't automatically inherit
  biproducts from $\cat{C}$
  \item however, $\PSh_{\CMon}(\cat{C})$ does indeed have biproducts if $\cat{C}$ does
  \item thus $\Fam(\PSh_{\CMon}(\FinVect))$ and $\Fam(\PSh_{\CMon}(\LatGal))$ are both Cartesian closed
  \item also need that $\CMon$-enriched Yoneda embedding preserves biproducts \todo{check: because this is how we will interpret base types? E.g.~$\sem{\mathsf{Bool}} = \Hom{\cat{C}}{-}{2}$ assuming $\cat{C}$ provides a suitable object 2}
  \end{enumerate}
}
